---
title: "Vedar Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vedar Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  echo = T, 
  message = F, 
  warning = F
)
```

```{r setup, echo = F}
library(vedar)
library(dplyr)
library(purrr)
library(igraph)
```

# Introduction

Veda is a graphical user interface for running TIMES models. Veda compiles model information specified in spreadsheets and passes the resulting data to GAMS and CPLEX for solving. The solution data is stored in custom veda text files. 

The Veda graphical user interface provides a means of interrogating the Veda solution output data. The vedar library offers an alternative and enables veda outputs to be imported and analysed in R.

# Getting veda data into R

The vignette uses the data that is created from processing demos model that were run in Veda2.0. 

```{r, warning = F}

filename_base <- "DemoS_001"

vd_filename <- paste(filename_base, ".vd", sep = "")
vds_filename <- paste(filename_base, ".vds", sep = "")
vde_filename <- paste(filename_base, ".vde", sep = "")
```


The vedar package has four functions used to load Veda model run output data into R.

The functions `import_vd(vd_filename)`, `import_vds(vds_filename)` and `import_vde(vde_filename)` files bring the veda run output data, the set information and the object descriptions respectively into the the workspace. In the code below, the `system_file()` is needed in the function arguments to access the vignette data, but this is not required when running in a regular project.

```{r, warning = F}

demos_vd <- import_vd(system.file("extdata", 
                                  vd_filename,
                                  package = "vedar"))

demos_vde <- import_vde(system.file("extdata", 
                                  vde_filename,
                                  package = "vedar"))

demos_vds <- import_vds(system.file("extdata", 
                                  vds_filename,
                                  package = "vedar"))

```

Alternatively, the three files can be imported and joined using `prep_data(filename_base)`. The `vignette = T` argument is used in the vignette, but is not needed in a project.

```{r, warning = F}
demos_001 <- prep_data(filename_base = filename_base, vignette = T)



```



# Assigning sectors 

Processes and commodities are likely to be assigned to particular grouping variables (e.g. sectors). A list of sector, or other grouping information can be associated to the data using `define_sector_from_list()`. As each row of the data has either or both commodity and process information, the sector is assigned by separate function calls on commodity and process data. 

Note that the joining variable can have different names in the veda output data, and the  sector information data. The former information is passed as a character (`join_variable_name`) and the latter as a column name (`sector_dat_join_variable_col`). The name of the grouping information is passed as a column (`sector_info_column`). 

The functionality of this will be improved in future iterations.


```{r, warning = F}
sectors_commodity_file <- "sectors_commodity.csv"

sectors_process_file <- "sectors_process.csv"
sectors_commodity <- system.file("extdata", 
                                  sectors_commodity_file,
                                  package = "vedar") %>%
  read.csv() %>%
  #convert to lower case strings
  prep_sector_dat()
sectors_process <- system.file("extdata", 
                                  sectors_process_file,
                                  package = "vedar") %>%
  read.csv() %>%
  #convert to lower case strings
  prep_sector_dat()

demos_001 <- demos_001 %>%
  #assign the sector to the commodity
  define_sector_from_list(join_variable_name = "commodity", 
                          sector_dat = sectors_commodity, 
                          sector_info_column = sector, 
                          sector_dat_join_variable_col = commodity) %>%
  #assign the sector to the process
  define_sector_from_list(join_variable_name = "process", 
                          sector_dat = sectors_process, 
                          sector_info_column = sector, 
                          sector_dat_join_variable_col = process) %>%
  # here, assign a sector based on the process sector if available,
  # else, use the commodity information
  mutate(sector = if_else(is.na(process_sector), 
                          commodity_sector, 
                          process_sector), 
         #create a variable to check that commodity sector and process
         # sector information match
         sector_match = if_else((is.na(process_sector) |
                                   is.na(commodity_sector) |
                                   process_sector == commodity_sector), 
                                T, F))
print(paste("Check if sector information matches between process and commodity: ", 
            sum(demos_001$sector_match) == nrow(demos_001), sep = ""))
```

# Comparing the results to veda output tables

We can compare results produced by vedar to those produced from tables generated by the Veda graphical-user interface
```{r}

# load results tables that were generated by Veda DemoS_001 example
data("demos_001_results")

flatten_dfr(demos_001_results["All costs.csv"]) %>%
  select(Attribute, 
         Commodity, 
         Process, 
         Period, 
         Region,
         Pv) %>%
  arrange(Attribute, Process, Period) %>%
  knitr::kable()


demos_001 %>% 
  #select rows in which attribute contains the string "cost_" 
  # and sector == "coal"
  filter(grepl("cost_", attribute), 
         sector == "coal") %>%
  select(attribute, 
         commodity, 
         process, 
         period, 
         region,
         pv) %>%
  arrange(attribute, process, period)  %>%
  knitr::kable()
```


# Representing the Reference Energy System as a graph object

Graphs/networks are representations of connected systems. The RES can be represented as a graph with processes as nodes, and commodities as edges.

`make_graph_from_veda_df()` converts the RES to an `igraph` object. When the data is for a single year, the edge weights are the value of the var_fout attribute for the given process/commodity.

```{r}
data(demos_001_sector)
g <- demos_001_sector %>%
      filter(period == 2006) %>%
      make_graph_from_veda_df(node_labels = process,
                              edge_labels = commodity
                              )




```

Converting the RES to a graph object enables the use of the network analysis functionality of the `igraph` package. For example, we can return all the edge weights


```{r}
print("Edge weights")
E(g)$weight
```

We can also compute all simple paths from a given node (or between two nodes if the `to = ` argument is specified)


```{r}

all_mincoa1_paths <- all_simple_paths(g, from = "mincoa1")
all_mincoa1_paths

```



# Viewing the Reference Energy System
The RES is a visual representation of processes and commodity flows. Vedar uses a Sankey diagram representation of the RES, with processes as nodes, and commodity flows as edges. The function `make_res()` creates the RES for a selected period, region and sector.


Here, we use the dataset from the DemoS_007 default run (DemoS_007). This is a multiregion model. The data has not had a sectors defined, so we first append sector variable for demonstration purposes. 

We first use the option of labeling nodes by the process_description. We label the edges by the commodity_description column, and set the font_size of the node labels.

Plotting the entire RES for 2020:

```{r, fig.width=7, , fig.height=5}
data("demos_007")

demos_007 <- demos_007 %>%
  mutate(sector = if_else(grepl("(oil)", process), "oil", "other"))

res_all <- make_res(dat = demos_007, 
                period_select = 2020, 
                node_labels = process_description, 
                edge_labels = commodity_description, 
                font_size = 10)

res_all



```
The RES can also be filtered for sectors and/or regions.

```{r, fig.width=7, fig.height=5}

res_filter <- make_res(dat = demos_007, 
                period_select = 2020, 
                sector_select = "oil", 
                region_select = "reg1",
                node_labels = process_description, 
                edge_labels = commodity_description, 
                font_size = 10)

res_filter

```

The flow labels are shown as tooltips when the mouse hovers over the edges. The magnitude of the flows is not included and all edges are shown with equal width. For visual representation, all demand commodities are shown to flow into a "demand commodity"_end_process.

To make the diagram less cluttered, the node label can use the process code. As an example, we select a different year and region for this example.

```{r, fig.width=7, fig.height=5}


res_prc_code <- make_res(dat = demos_007, 
                period_select = 2015, 
                region_select = "reg1", 
                sector_select = "oil", 
                node_labels = process, 
                edge_labels = commodity_description, 
                font_size = 10)

res_prc_code



```
