---
title: "Vedar Vigenette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vedar Vigenette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(vedar)
```

# Introduction


# Getting veda data into R
The vignette uses the data that is created from processing the demos_001 model that is run in Veda2.0. 

The vedar package has four functions used to load Veda model run output data into R.

The functions `import_vd()`, `import_vds()` and `import_vde()` files bring the veda run output data, the set information and the object descriptions respectively into the the workspace. In the example 

```{r, warning = F, echo=T, eval = F}

filename_base <- "DemoS_001"

vd_filename <- paste(filename_base, ".vd", sep = "")
vds_filename <- paste(filename_base, ".vds", sep = "")
vde_filename <- paste(filename_base, ".vde", sep = "")

demos_vd <- import_vd(vd_filename)

demos_vde <- import_vd(vds_filename)

demos_vds <- import_vd(vde_filename)

```

```{r, warning = F, echo=F}

filename_base <- "DemoS_001"

vd_filename <- paste(filename_base, ".vd", sep = "")
vds_filename <- paste(filename_base, ".vds", sep = "")
vde_filename <- paste(filename_base, ".vde", sep = "")

demos_vd <- import_vd(system.file("extdata", 
                                  vd_filename,
                                  package = "vedar"))

demos_vde <- import_vd(system.file("extdata", 
                                  vde_filename,
                                  package = "vedar"))

demos_vds <- import_vd(system.file("extdata", 
                                  vds_filename,
                                  package = "vedar"))

```

Alternatively, the three files can be imported and joined using `prep_data()`

```{r, warning = F, eval = F}
demos_001 <- prep_data(filename_base = filename_base)



```


```{r, warning = F, eval = T, echo = F}
demos_001 <- prep_data_for_vignette(filename_base = filename_base)



```
